A Logical Solution of the Linearity Challenge

### 2025-12-02 paper/Permutations.lagda.tex
cleanup

This definition is essentially the same as the notion of list permutation in
Agda's standard library. Each constructor represents a particular kind of
permutation: \AgdaInductiveConstructor{refl} for the trivial permutation that
does not change anything; \AgdaInductiveConstructor{swap} for the permutation
that swaps the first two elements of a typing context;
\AgdaInductiveConstructor{prep} for the permutation that applies a permutation
to the tail of a typing context; \AgdaInductiveConstructor{trans} for the
sequential composition of permutations.
%
We have preferred defining our own notion of permutation for simplicity: the
\AgdaInductiveConstructor{swap} constructor does not need a sub-permutation for
the tail of the typing context, which can always be performed, if needed, using
\AgdaInductiveConstructor{prep} and \AgdaInductiveConstructor{trans}. Also, and
more importantly, this definition is \emph{monomorphic} (it does not need to
operate on lists of arbitrary type), resulting in a more streamlined usage in
the rest of the code.

@@

Just like permutations from Agda's standard library, also \AgdaDatatype{\_↭\_}
enjoys a number of expected (and useful) properties, such as symmetry. Among all
properties, in the following we will make use of a specific one that connects
context splitting and permutations: if $\Splitting\ContextC\ContextD\ContextE$,
then $\ContextC$ is a permutation of the concatenation of $\ContextD$ and
$\ContextE$.\Luca{Ci sono molte altre proprietà non descritte, controllare
quelle che servono davvero.}

>>>>

Each constructor of \AgdaDatatype{\_↭\_} represents a particular kind of
permutation: \AgdaInductiveConstructor{refl} for the trivial permutation that
does not change anything; \AgdaInductiveConstructor{swap} for the permutation
that swaps the first two elements of a typing context;
\AgdaInductiveConstructor{prep} for the permutation applied to the tail of a
typing context; \AgdaInductiveConstructor{trans} for the sequential composition
of permutations.

The definition of \AgdaDatatype{\_↭\_} data type is nearly the same found in
Agda's standard library. We have preferred defining our own notion of
permutation for improved simplicity and convenience: the
\AgdaInductiveConstructor{swap} constructor does not need a sub-permutation for
the tail of the typing context, which can always be performed, if needed, using
\AgdaInductiveConstructor{prep} and \AgdaInductiveConstructor{trans}. Also, and
more importantly, the data type is \emph{monomorphic} (it does not need to
relate lists of arbitrary type) and the quantified arguments $A$ and $B$ of
\AgdaInductiveConstructor{swap} and \AgdaInductiveConstructor{prep} are
implicit, which streamlines their usage in the rest of the code.

@@

Just like permutations from Agda's standard library, also \AgdaDatatype{\_↭\_}
enjoys a number of expected properties, such as symmetry. Among them, in the
following we will make use of a specific one that connects context splitting and
permutations: if $\Splitting\ContextC\ContextD\ContextE$, then $\ContextC$ is a
permutation of the concatenation of $\ContextD$ and $\ContextE$.\Luca{Ci sono
molte altre proprietà non descritte, controllare quelle che servono davvero.}
